The human in the loop: volunteer-based metacomputers as a socio-technical system

### 2016-04-24 Alife16/paper.tex
Including a definition of the cache
Which was there, only not too clear. Closes #111

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the `CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \citep{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. This data structure can
  hold every chromosome in a particular experiment, or have a finite
  size that erases the oldest chromosomes once it has filled to
  capacity. In this paper we will test both implementations.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client has done. 
\end{enumerate}

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS \url{www.openshift.com}, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts on Twitter and, in the latest case, Telegram, and
results were published in\citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February. All  %% April --> January ??
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
%Pablo: but it would great at least describe how many times was announced and how long
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions resets to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that. In fact,
the islands running in the browser {\em spill} from one experiment to
the next. 
If an individual that is close or even at optimum value is stored in the cache 
the next experiment will be affected, and it will require less time 
to finish. This problem has been addressed in later versions of the
system, however, time to solution is not the important measure here,
where we are concerned with the computational 
power of the system established by the number of volunteers.

The table shows that every experiment included more than 30 runs. The
%Pablo: reference to table? (maybe the final print will misplace it)
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment, with a
number around 50 in the second, and most recent, batch of experiments. 
%
\begin{table}[htb]
\caption{Summary of time per run, number of IPs and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|cccc}
\hline
     & \multicolumn{2}{c}{IPs} & \multicolumn{2}{c}{Median} \\
Experiment & Median & Max & time (s) &  \#{\tt  PUT}s \\
\hline
4/4 & 5 & 16 & 2040 & 18   \\
4/24 &  5 & 29 & 732 & 11  \\
7/31 & 5 & 14 & 260 & 23   \\
\hline
Cache=128 & 5 & 17 & 222.2 & 124 \\ 
Cache=64 & 8 & 38 & 51.3 & 100 \\
Cache=32 & 6 & 19 & 58.9 & 45 \\
\hline
% Data in .RData file in Splash-volunteer
\end{tabular}
\end{center}
\end{table}
%

>>>>

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the `CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \citep{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a key-value store
  that is reset when the solution is found. This store can
  hold every chromosome in a particular experiment, or have a finite
  size that erases the oldest chromosomes once it has filled to
  capacity, acting as a cache. In this paper we will test both implementations.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client has done. 
\end{enumerate}

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS \url{www.openshift.com}, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts on Twitter and, in the latest case, Telegram, and
results were published in\citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February of the next year. All  
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
%Pablo: but it would great at least describe how many times was
%announced and how long
% We don't really tracked this... - JJ
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions resets to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that. In fact,
the islands running in the browser {\em spill} from one experiment to
the next. 
If an individual that is close or even at optimum value is stored in the cache 
the next experiment will be affected, and it will require less time 
to finish. This problem has been addressed in later versions of the
system, however, time to solution is not the important measure here,
where we are concerned with the computational 
power of the system established by the number of volunteers.

The table \ref{tab:runs}shows that every experiment included more than 30 runs. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment, with a
number around 50 in the second, and most recent, batch of experiments. 
%
\begin{table}[htb]
\caption{Summary of time per run, number of IPs and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|cccc}
\hline
     & \multicolumn{2}{c}{IPs} & \multicolumn{2}{c}{Median} \\
Experiment & Median & Max & time (s) &  \#{\tt  PUT}s \\
\hline
4/4 & 5 & 16 & 2040 & 18   \\
4/24 &  5 & 29 & 732 & 11  \\
7/31 & 5 & 14 & 260 & 23   \\
\hline
Cache=128 & 5 & 17 & 222.2 & 124 \\ 
Cache=64 & 8 & 38 & 51.3 & 100 \\
Cache=32 & 6 & 19 & 58.9 & 45 \\
\hline
% Data in .RData file in Splash-volunteer
\end{tabular}
\end{center}
\end{table}
%
